Intentional release of native species undermines ecological stability

### 2022-03-12 rmd/figures.Rmd
Squashed commit of the following:
commit 42b7838e55c1587f6a744b0a978c367eac453d43
Author: AkiraTerui <hanabi0111@gmail.com>
Date:   Sat Mar 12 11:50:40 2022 -0500

    edit: affiliation

commit 1933aaca4b7cc7f4d29771b59b45cf512e403b48
Author: aterui <hanabi0111@gmail.com>
Date:   Thu Mar 10 22:06:24 2022 -0500

    edit:  manuscript

commit 0884ad16cc1eb71df364fe1aaf0e7cec0dadd794
Author: aterui <hanabi0111@gmail.com>
Date:   Thu Mar 10 21:32:32 2022 -0500

    edit: manuscript

commit db4e8b6bab9c85a4ba5ba4767ac5442fcf493553
Author: AkiraTerui <hanabi0111@gmail.com>
Date:   Thu Mar 10 19:04:2

**Figure 1** Theory and empirical observations agree that stock enhancement destabilizes community dynamics. In A and B, panels and colors distinguish response variables and species groups. (A) Theoretical predictions. Dots represent individual simulation replicates, and lines and shades are loess curves fitted to simulated data and 95% confidence intervals. Parameters used in this simulation are: intrinsic growth rate of an enhanced species $r_1=$ `r unique(df0$r1)`, average interspecific competition $\bar{\alpha}=$ `r unique(df0$alpha)`, carrying capacity $K=$ `r unique(df0$k)`, environmental stochasticity $\sigma_{\epsilon}=$ `r unique(df0$sd_env)`, relative fitness of captive-bred individuals $f_R=$ `r unique(df0$phi)`. (B) Empirical evidence. Dots represent geometric means of site-level observations in 31 watersheds. Lines are the predicted values of the regression models, and line types correspond to the coefficient's posterior probabilities. See Tables S8-10 for full statistics. (C) Map of sampling sites (black dots) in Hokkaido, Japan. (D) Example of a protected watershed (Okushibetsu). (E) Masu salmon *Oncorhynchus masou masou*. Photo credit: Akira Terui.

@@

**Figure 2** Posterior distributions for the standardized regression coefficients of stock enhancement. Y-axis represents different response variables grouped by colors distinguishing species groups (whole community, enhanced species [masu salmon], and unenhanced species).

>>>>

**Figure 1** Theory and empirical observations agree that intentional release destabilizes community dynamics. In A and B, panels and colors distinguish response variables and species groups. (A) Theoretical predictions. Dots represent individual simulation replicates, and lines and shades are loess curves fitted to simulated data and 95% confidence intervals. Parameters used in this simulation are: intrinsic growth rate of an enhanced species $r_1=$ `r unique(df0$r1)`, average interspecific competition $\bar{\alpha}=$ `r unique(df0$alpha)`, carrying capacity $K=$ `r unique(df0$k)`, environmental stochasticity $\sigma_{\epsilon}=$ `r unique(df0$sd_env)`, relative fitness of captive-bred individuals $f_R=$ `r unique(df0$phi)`. (B) Empirical evidence. Dots represent geometric means of site-level observations in 31 watersheds. Lines are the predicted values of the regression models, and line types correspond to the coefficient's posterior probabilities. See Tables S8-10 for full statistics. (C) Map of sampling sites (black dots) in Hokkaido, Japan. (D) Example of a protected watershed (Okushibetsu). (E) Masu salmon *Oncorhynchus masou masou*. Photo credit: Akira Terui.

@@

**Figure 2** Posterior distributions for the standardized regression coefficients of intentional release. Y-axis represents different response variables grouped by colors distinguishing species groups (whole community, enhanced species [masu salmon], and unenhanced species).

### 2022-03-12 rmd/manuscript.Rmd
Squashed commit of the following:
commit 42b7838e55c1587f6a744b0a978c367eac453d43
Author: AkiraTerui <hanabi0111@gmail.com>
Date:   Sat Mar 12 11:50:40 2022 -0500

    edit: affiliation

commit 1933aaca4b7cc7f4d29771b59b45cf512e403b48
Author: aterui <hanabi0111@gmail.com>
Date:   Thu Mar 10 22:06:24 2022 -0500

    edit:  manuscript

commit 0884ad16cc1eb71df364fe1aaf0e7cec0dadd794
Author: aterui <hanabi0111@gmail.com>
Date:   Thu Mar 10 21:32:32 2022 -0500

    edit: manuscript

commit db4e8b6bab9c85a4ba5ba4767ac5442fcf493553
Author: AkiraTerui <hanabi0111@gmail.com>
Date:   Thu Mar 10 19:04:2

---
title: "Intentional stocking undermines ecological stability"
output:
  #pdf_document:
  #  latex_engine: xelatex
  word_document:
    reference_docx: word_format.docx
csl: science.csl
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
bibliography: reference.bib
fontsize: 11pt
header-includes:
  - \usepackage{lineno}
  - \linenumbers
---

@@

## stocking data
source(here::here("code/data_fmt_stock.R"))

@@

**Authors:** Akira Terui^1,\*^, Hirokazu Urabe^2^, Masayuki Senzaki^3^, Bungo Nishizawa^4^

\*Corresponding author: hanabi0111\@gmail.com

**Affiliations:** ^1^ Department of Biology, University of North Carolina at Greensboro, Greensboro, NC 27412, USA ^2^ Salmon and Freshwater Fisheries Research Institute, Hokkaido Research Organization, Eniwa, Hokkaido, 061-1433, Japan ^3^ Faculty of Environmental Earth Science, Hokkaido University, Sapporo, Hokkaido, 060-0810, Japan ^4^ National Institute of Polar Research, Tachikawa, Tokyo, 190-8518, Japan

@@

The past decades have witnessed efforts to unveil ecological risks associated with massive releases of captive-bred individuals ("stock enhancement"). Yet, this method is still pervasive in conservation and natural resource management, assuming that the demographic benefits exceed the associated risks. Here, contrary to this intuition, we show that stock enhancement triggers community collapse with little demographic benefit to the enhanced species. Our theory predicted that stock enhancement destabilizes community dynamics by facilitating competitive exclusion. In parallel, the excessive input of captive-bred individuals suppressed the natural recruitment of the enhanced species via intensified within-species competition. Consistent with this prediction, stream fish communities showed greater temporal fluctuations and fewer species richness in rivers with the intensive stocking of hatchery masu salmon - a major freshwater resource in Japan. Our findings alarm that the current overreliance on stock enhancement may accelerate global biodiversity loss, threatening the provisioning of ecosystem services. A radical change in management strategies is therefore required for the sustainability of natural resources.

@@

Human demands for natural resources are ever-increasing, such that active interventions are critical to the sustainable management of fisheries, forestry, and wildlife [@laikreCompromisingGeneticDiversity2010]. Captive breeding is a form of the efforts to enhance wild populations of diverse plant and animal taxa [@snyderLimitationsCaptiveBreeding1996; @laikreCompromisingGeneticDiversity2010]. Although releases of captive-bred organisms ("stock enhancement") entail ecological risks such as the accumulation of deleterious alleles [@arakiGeneticEffectsCaptive2007; @krkosekDecliningWildSalmon2007], this method is still pervasive in conservation [@snyderLimitationsCaptiveBreeding1996] and natural resource management [@laikreCompromisingGeneticDiversity2010]. In fisheries, for example, billions of hatchery individuals are released annually across the globe [@kitadaEconomicEcologicalGenetic2018]. The widespread use of captive breeding is perhaps because of the "myth" that the demographic benefits of released individuals may exceed the associated risks.

Current debates, however, overlook the fact that we have rarely assessed the community-wide impact of stock enhancement. Species are all embedded in the complex web of interacting organisms, and the stable coexistence of competing species through density-dependent feedback underpins the emergent stability of ecological communities [@tilmanEcologicalConsequencesChanges1999; @carrollNicheFitnessDifferences2011a; @loreauDoesFunctionalRedundancy2004]. Stock enhancement may disrupt the sensitive balance of species interactions because it introduces unnaturally high numbers of individuals into the wild [@careyNativeInvadersChallenges2012]. Hence, this form of species management may intervene in the ecological process that allows competing species to coexist, ultimately degrading the long-term community stability. Evidence for this hypothesis is lacking, however.

Here, we show that stock enhancement undermines long-term community stability, which we define as the relative size of fluctuations in total community density over time [@tilmanEcologicalConsequencesChanges1999]. Our theory illuminates that stock enhancement compromises the stabilizing mechanism emerging from species niche differences. The present study further demonstrates the relevance of this general theory to natural systems by showing its congruence with Japanese stream fish communities, where \~10 million hatchery masu salmon (*Oncorhynchus masou masou*) are released annually for fisheries, recreation, and conservation purposes [@kitadaLessonsJapanMarine2020].

We employed a multispecies Ricker model [@fowlerSpeciesDynamicsAlter2012] to simulate community dynamics with the selective stock enhancement of a constituent species (species 1). Specifically, the population density of species $i$ at time $t+1$, $N_{i,t+1}$, is modeled as:

@@

We ran `r n_sim` time steps of `r max(sim_result$n_rep)` independent communities (i.e., simulation replicates) under each of `r nrow(param_set)` simulation scenarios. These scenarios cover a range of ecological contexts, differing in the intrinsic growth of an enhanced species $r_1$, competition $\bar{\alpha}$, carrying capacity $K$, and relative fitness $f_R$ (see **Methods**). Using the last `r unique(sim_result$n_timestep)` time steps, we obtained the following summary statistics of the total community density $\sum_i^SN_i$ to examine the community-level response to stock enhancement: coefficient of variation (CV), number of species persist (defined as $N_i > 0.01$ at $t=1600$), temporal mean ($\mu$), and temporal SD ($\sigma$). We also calculated the temporal mean and SD for the enhanced ($N_1$) and unenhanced species ($\sum_{i,i \ne 1}^S N_i$) separately to infer underlying mechanisms.

When carrying capacity was small ($K=100$), our model predicted a destabilizing effect of stock enhancement on ecological communities as illustrated by increased CV with increasing numbers of releases (**Figure 1A**). This pattern stemmed mainly from the reduced mean of the total community density, and both enhanced and unenhanced species groups were responsible (**Figure 1A**). The enhanced species decreased because stocking induced the negative competitive effect exceeding the reproductive contribution of released individuals [@satakeStockingCaptivebredFish2012]. Meanwhile, interspecific competition reduced the unenhanced species at high levels of stock enhancement, resulting in fewer persisting species (**Figure 1A**). Combined, the total community density decreased more sharply than individual species groups (**Figure 1A**). The SDs showed a similar trend, but the relationship was flatter at the community level (**Figure 1A**). Since a CV is a ratio of an SD to a mean, the steeper decline of the mean community density led to the increased CV. These patterns were qualitatively similar across most ecological contexts (robust to the changes in $r_1$, $\bar{\alpha}$, and $f_R$; **Figures S1-S3**).

The destabilizing effect emerges because stock enhancement affects the balance of species interactions that underpins community stability. In theory, the stable coexistence requires a niche difference that is large enough to overcome the relative difference in intrinsic competitive ability [@carrollNicheFitnessDifferences2011a]. Under this condition, competing species can grow from small populations because dominant species undergo stronger intraspecific competition [@carrollNicheFitnessDifferences2011a]. Such coexistence favors stable temporal dynamics of species-rich communities [@tilmanEcologicalConsequencesChanges1999] because it gives rise to "overyielding" [@loreauDoesFunctionalRedundancy2004; @carrollNicheFitnessDifferences2011a], i.e., total community density of a multispecies community exceeds what would be expected in a single species community ($\sum_i^S N_i > K$). However, stock enhancement is externally controlled, and the number of releases is not subject to density-dependent regulation. Therefore, released individuals impose additional intra- and interspecific competition that interferes with the ecological process producing overyielding.

Stock enhancement, however, had little influence on community dynamics when carrying capacity was sufficiently large ($K=500$; **Figures S4-S6**). In particular, stock enhancement increased the enhanced species with a low population growth rate ($r_1=0.5$; **Figure S4**). This result may explain why some of the best evidence for successful stock enhancement comes from long-lived endangered species [e.g., @jonesRestorationMauritiusKestrel1995]. More importantly, the contrasting community response at different carrying capacities provides deeper insights. One of the core motivations for stock enhancement is to mitigate declining trends of natural resources due to human impacts, such as habitat loss [@chaseEcosystemDecayExacerbates2020]. Ironically, our results suggest that stock enhancement will not bring desired outcomes unless we resolve root causes that compromise environmental capacity first.

To demonstrate the relevance of our general theory to natural systems, we assessed the potential impacts of the stock enhancement of masu salmon (**Figure 1E**) on the long-term stability of stream fish communities in Hokkaido, Japan. In the protected watersheds (all separated by the ocean; **Figure 1C, D**), a long-term program exists to monitor stream fish communities along with the official stocking records. The majority of stocking occurs in spring, after which salmon fry stay in freshwater for growing seasons. Therefore, the study system sets the stage for a "natural experiment" to test our theoretical predictions. We used the data from 1999 to 2019 at `r n_distinct(df_fish$site_id)` sites within `r n_distinct(df_fish$river)` independent watersheds (see **Methods** for selection criteria). Using hierarchical Bayesian models, we quantified the effect of stock enhancement on community dynamics while accounting for potential effects of climates and local abiotic factors.

As predicted, stream fish communities showed greater temporal fluctuations (higher CV) in watersheds with intensive stocking (**Figure 1B**). The effect was striking in its magnitude, almost doubling the CV at the highest stocking level. Our analysis strongly supported the positive relationship between the CV and the number of releases (**Figure 2**), in which the probability of the regression coefficient being positive was `r op(df_beta$all$prob_positive[df_beta$all$response == "cv"])` (**Table S8**). This pattern was associated with the reduced long-term average of the total community density and fewer species richness (**Figures 1B and 2**), and both enhanced (masu salmon) and unenhanced fish groups contributed to this trend (**Figures 1B and 2;** see **Tables S8-S10** for full statistics). In the meantime, the SDs had vague relationships with stock enhancement (**Figures 1B and 2**).

@@

We should prioritize habitat conservation with a broader scope of ecosystem management for the sustainability of natural resources. Such efforts are particularly important in the Anthropocene because habitat degradation may exacerbate the undesired influence of stock enhancement. Protected areas and environmental restoration are promising tools to conserve biodiversity, and a smart spatial design is a key to achieving successful conservation. For example, coordinated placement of conservation sites considering spatial biodiversity patterns is crucial in improving the ecological outcomes [@teruiEmergentDualScaling2021; @koningNetworkGrassrootsReserves2020; @fontouraProtectingConnectivityPromotes2022]. Governance may also play a central role in enforcing environmental legislation, potentially determining the effectiveness of conservation investment [@amanoSuccessfulConservationGlobal2018]. These considerable potentials indicate that viable management options exist before blindly accepting stock enhancement. Without a comprehensive framework that appreciates the ecological integrity of natural communities, the stock enhancement will never be effective but impairs biodiversity.

>>>>

---
title: "Intentional release of native species undermines ecological stability"
output:
  pdf_document:
    latex_engine: xelatex
  #word_document:
  # reference_docx: word_format.docx
csl: nature.csl
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
bibliography: reference.bib
fontsize: 11pt
header-includes:
  - \usepackage{lineno}
  - \linenumbers
---

@@

## release data
source(here::here("code/data_fmt_stock.R"))

@@

@@

The past decades have witnessed efforts to unveil ecological risks associated with massive releases of captive-bred native species ("intentional release"). Yet, this method is still pervasive in conservation and natural resource management, assuming that the demographic benefits exceed the associated risks. Here, contrary to this intuition, we show that intentional release triggers community collapse with little demographic benefit to the enhanced species. Our theory predicted that intentional release destabilizes community dynamics by facilitating competitive exclusion. In parallel, the excessive input of captive-bred individuals suppressed the natural recruitment of the enhanced species via intensified within-species competition. Consistent with this prediction, stream fish communities showed greater temporal fluctuations and fewer species richness in rivers with the intensive release of hatchery masu salmon - a major freshwater resource in Japan. Our findings alarm that the current overreliance on intentional release may accelerate global biodiversity loss, threatening the provisioning of ecosystem services.

@@

Human demands for natural resources are ever-increasing, such that active interventions are critical to the sustainable management of fisheries, forestry, and wildlife [@laikreCompromisingGeneticDiversity2010]. Captive breeding is a form of the efforts to enhance wild populations of diverse plant and animal taxa [@snyderLimitationsCaptiveBreeding1996; @laikreCompromisingGeneticDiversity2010]. Although releases of captive-bred native species ("intentional release") entail ecological risks such as the accumulation of deleterious alleles [@arakiGeneticEffectsCaptive2007; @krkosekDecliningWildSalmon2007], this method is still pervasive in conservation [@snyderLimitationsCaptiveBreeding1996] and natural resource management [@laikreCompromisingGeneticDiversity2010]. In fisheries, for example, billions of hatchery individuals are released annually across the globe [@kitadaEconomicEcologicalGenetic2018]. The widespread use of captive breeding is perhaps because of the "myth" that the demographic benefits of released individuals may exceed the associated risks.

Current debates, however, overlook the fact that we have rarely assessed the community-wide impact of intentional release. Species are all embedded in the complex web of interacting organisms, and the stable coexistence of competing species through density-dependent feedback underpins the emergent stability of ecological communities [@tilmanEcologicalConsequencesChanges1999; @carrollNicheFitnessDifferences2011a; @loreauDoesFunctionalRedundancy2004]. Intentional release may disrupt the sensitive balance of species interactions because it introduces unnaturally high numbers of individuals into the wild [@careyNativeInvadersChallenges2012]. Hence, this form of species management may intervene in the ecological process that allows competing species to coexist, ultimately degrading the long-term community stability. Evidence for this hypothesis is lacking, however.

Here, we show that intentional release undermines long-term community stability, which we define as the relative size of fluctuations in total community density over time [@tilmanEcologicalConsequencesChanges1999]. Our theory illuminates that intentional release compromises the stabilizing mechanism emerging from species niche differences. The present study further demonstrates the relevance of this general theory to natural systems by showing its congruence with Japanese stream fish communities, where \~10 million hatchery masu salmon (*Oncorhynchus masou masou*) are released annually for fisheries, recreation, and conservation purposes [@kitadaLessonsJapanMarine2020].

We employed a multispecies Ricker model [@fowlerSpeciesDynamicsAlter2012] to simulate community dynamics with the selective intentional release of a constituent species (species 1). Specifically, the population density of species $i$ at time $t+1$, $N_{i,t+1}$, is modeled as:

@@

We ran `r n_sim` time steps of `r max(sim_result$n_rep)` independent communities (i.e., simulation replicates) under each of `r nrow(param_set)` simulation scenarios. These scenarios cover a range of ecological contexts, differing in the intrinsic growth of an enhanced species $r_1$, competition $\bar{\alpha}$, carrying capacity $K$, and relative fitness $f_R$ (see **Methods**). Using the last `r unique(sim_result$n_timestep)` time steps, we obtained the following summary statistics of the total community density $\sum_i^SN_i$ to examine the community-level response to intentional release: coefficient of variation (CV), number of species persist (defined as $N_i > 0.01$ at $t=1600$), temporal mean ($\mu$), and temporal SD ($\sigma$). We also calculated the temporal mean and SD for the enhanced ($N_1$) and unenhanced species ($\sum_{i,i \ne 1}^S N_i$) separately to infer underlying mechanisms.

When carrying capacity was small ($K=100$), our model predicted a destabilizing effect of intentional release on ecological communities as illustrated by increased CV with increasing numbers of releases (**Figure 1A**). This pattern stemmed mainly from the reduced mean of the total community density, and both enhanced and unenhanced species groups were responsible (**Figure 1A**). The enhanced species decreased because release induced the negative competitive effect exceeding the reproductive contribution of released individuals [@satakeStockingCaptivebredFish2012]. Meanwhile, interspecific competition reduced the unenhanced species at high levels of intentional release, resulting in fewer persisting species (**Figure 1A**). Combined, the total community density decreased more sharply than individual species groups (**Figure 1A**). The SDs showed a similar trend, but the relationship was flatter at the community level (**Figure 1A**). Since a CV is a ratio of an SD to a mean, the steeper decline of the mean community density led to the increased CV. These patterns were qualitatively similar across most ecological contexts (robust to the changes in $r_1$, $\bar{\alpha}$, and $f_R$; **Figures S1-S3**).

The destabilizing effect emerges because intentional release affects the balance of species interactions that underpins community stability. In theory, the stable coexistence requires a niche difference that is large enough to overcome the relative difference in intrinsic competitive ability [@carrollNicheFitnessDifferences2011a]. Under this condition, competing species can grow from small populations because dominant species undergo stronger intraspecific competition [@carrollNicheFitnessDifferences2011a]. Such coexistence favors stable temporal dynamics of species-rich communities [@tilmanEcologicalConsequencesChanges1999] because it gives rise to "overyielding" [@loreauDoesFunctionalRedundancy2004; @carrollNicheFitnessDifferences2011a], i.e., total community density of a multispecies community exceeds what would be expected in a single species community ($\sum_i^S N_i > K$). However, intentional release is externally controlled, and the number of releases is not subject to density-dependent regulation. Therefore, released individuals impose additional intra- and interspecific competition that interferes with the ecological process producing overyielding.

Intentional release, however, had little influence on community dynamics when carrying capacity was sufficiently large ($K=500$; **Figures S4-S6**). In particular, intentional release increased the enhanced species with a low population growth rate ($r_1=0.5$; **Figure S4**). This result may explain why some of the best evidence for successful intentional release comes from long-lived endangered species [e.g., @jonesRestorationMauritiusKestrel1995]. More importantly, the contrasting community response at different carrying capacities provides deeper insights. One of the core motivations for intentional release is to mitigate declining trends of natural resources due to human impacts, such as habitat loss [@chaseEcosystemDecayExacerbates2020]. Ironically, our results suggest that intentional release will not bring desired outcomes unless we resolve root causes that compromise environmental capacity first.

To demonstrate the relevance of our general theory to natural systems, we assessed the potential impacts of the intentional release of masu salmon (**Figure 1E**) on the long-term stability of stream fish communities in Hokkaido, Japan. In the protected watersheds (all separated by the ocean; **Figure 1C, D**), a long-term program exists to monitor stream fish communities along with the official release records. The majority of release occurs in spring, after which salmon fry stay in freshwater for growing seasons. Therefore, the study system sets the stage for a "natural experiment" to test our theoretical predictions. We used the data from 1999 to 2019 at `r n_distinct(df_fish$site_id)` sites within `r n_distinct(df_fish$river)` independent watersheds (see **Methods** for selection criteria). Using hierarchical Bayesian models, we quantified the effect of intentional release on community dynamics while accounting for potential effects of climates and local abiotic factors.

As predicted, stream fish communities showed greater temporal fluctuations (higher CV) in watersheds with intensive release (**Figure 1B**). The effect was striking in its magnitude, almost doubling the CV at the highest release level. Our analysis strongly supported the positive relationship between the CV and the number of releases (**Figure 2**), in which the probability of the regression coefficient being positive was `r op(df_beta$all$prob_positive[df_beta$all$response == "cv"])` (**Table S8**). This pattern was associated with the reduced long-term average of the total community density and fewer species richness (**Figures 1B and 2**), and both enhanced (masu salmon) and unenhanced fish groups contributed to this trend (**Figures 1B and 2;** see **Tables S8-S10** for full statistics). In the meantime, the SDs had vague relationships with intentional release (**Figures 1B and 2**).

@@

We should prioritize habitat conservation with a broader scope of ecosystem management for the sustainability of natural resources. Such efforts are particularly important in the Anthropocene because habitat degradation may exacerbate the undesired influence of intentional release. Protected areas and environmental restoration are promising tools to conserve biodiversity, and a smart spatial design is a key to achieving successful conservation. For example, coordinated placement of conservation sites considering spatial biodiversity patterns is crucial in improving the ecological outcomes [@teruiEmergentDualScaling2021; @koningNetworkGrassrootsReserves2020; @fontouraProtectingConnectivityPromotes2022]. Governance may also play a central role in enforcing environmental legislation, potentially determining the effectiveness of conservation investment [@amanoSuccessfulConservationGlobal2018]. These considerable potentials indicate that viable management options exist before blindly accepting intentional release. Without a comprehensive framework that appreciates the ecological integrity of natural communities, the intentional release will never be effective but impairs biodiversity.
